---
title: 'Specification analysis for technology use and teenager well-being'
subtitle: 'SCA and BSCA estimator properties: a simulation study'
author: 
  - Christoph Semken
  - David Rossell
bibliography: specurveanalysis.bib
output:
  html_document:
    toc: yes
    number_sections: yes
---

```{r setup sim, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(warn=-1, scipen=999)
options(xtable.table.placement="bt")
```

This file contains the simulation to compare the properties of SCA and BMA estimators. We obtain the following results:

* Scenario 1 ($\beta_1=0$): ``r knitr::load_cache('tabsummary', 'tab1summary')``
* Scenario 2 ($\beta_1=1$): ``r knitr::load_cache('tabsummary', 'tab2summary')``
* Scenario 3 ($\beta_1=\beta_2=\beta_3=\beta_4=\beta_5=\beta_6=0$): ``r knitr::load_cache('tabsummary', 'tab3summary')``
* Scenario 4 ($\beta_1=\beta_2=0,\beta_3=0.25,\beta_4=0.75,\beta_5=\beta_6=1$): ``r knitr::load_cache('tabsummary', 'tab4summary')``

Subsection 1 sets up some useful functions to run the simulations.
Subsection 2 considers a simplest setting with a single treatment variable of interest. BMA and the SCA median are used to average inference across models, that is across possible specifications defined by each treatment/control having/not having an effect. 
Subsection 3 considers multiple treatments of interest. We show how to obtain inference for all individual treatment effects, so that one can assess and report on their heterogeneity. We also show how to estimate and test for the presence on an average treatment effect.
Finally, Subsection 4 studies the properties of the BMA estimator in the case the ATE is taken over multiple outcomes.


# Setup

We use the packages `mombf` and `specr` for BMA and SCA inference, respectively.

```{r setup sim imports, message=FALSE}
library(mombf)
library(specr)
library(tidyverse)
library(xtable)
```

We also load our auxiliary functions.

```{r,eval=FALSE}
source('code/functions.R')
```

```{r setup sim notshow, echo=FALSE}
PATH = Sys.getenv("BSCA_PATH")

source(file.path(PATH,'code/functions.R'))
```

We create a function that simulates data from a Gaussian linear regression with $n$ observations: the outcome variable $y$, treatment(s) $x$ and control covariates $z$. The true value of the regression parameters for $x$ and $z$ are specified in $\beta$ and $\alpha$, respectively. The treatment(s) and controls are correlated, i.e. this is a situation where it is necessary to identify the relevant controls to avoid parameter estimation bias and increase the power of statistical tests.

```{r}
simdata= function(n, beta, alpha, seed) {
 set.seed(seed)
 J= length(beta); q= length(alpha)
 x= matrix(rnorm(n*J),nrow=n) #simulate the value of treatment(s)
 xnames= paste('x',1:length(beta),sep='')
 if (q>0) { #if there are control covariates
   z = rowMeans(x) + matrix(rnorm(n*q),nrow=n) #correlated with treatments
   y = x %*% matrix(beta,ncol=1) + z %*% matrix(alpha,ncol=1) + rnorm(n)
   znames= paste('z',1:length(alpha),sep='')
   ans= data.frame(y,1,x,z)
   colnames(ans)= c('y','Intercept',xnames,znames)
 } else {   #if there are no control covariates
   y = x %*% matrix(beta,ncol=1) + rnorm(n)
   ans= data.frame(y,1,x)
   colnames(ans)= c('y','Intercept',xnames)
 }
 return(ans)
}
```


# Single treatment

## Truly no treatment effect

BMA point estimates are stored in `bmaest` and its posterior probabilities on the presence of an effect in `margpp`. We also store in `scaest` the point estimate returned by a standard Specification Curve Analysis, which takes the median across all possible specifications. 
For the SCA median we use the paired bootstrap procedure proposed by Simonsohn et al. (2020).

```{r}
n= 1000; beta= 0; alpha= 1; nsims= 100; nperm= 500; 
xnames= paste('x',1:length(beta),sep='')
znames= paste('z',1:length(alpha),sep='')
bmaest= margpp= matrix(NA,nrow=nsims,ncol=1+length(beta)+length(alpha))
colnames(bmaest)= colnames(margpp)= c('Intercept', xnames, znames)
scaest= scapval= double(nsims)
for (i in 1:nsims) {
    data= simdata(n=n,beta=beta,alpha=alpha,seed=300+i)
    #BMA
    ms = modelSelection(
      y=data[,1], x=data[,-1],verbose=FALSE,
      priorCoef=zellnerprior(tau=nrow(data))
    )
    b= coef(ms)
    bmaest[i,]= b[-nrow(b),'estimate']
    margpp[i,]= b[-nrow(b),'margpp']
    # SCA median bootstrap-based permutation test 
    # "(1) Estimate all K specifications with the observed data"
    sca= run_specs(
      df=data, y="y", x=xnames, controls=znames, 
      model="lm", keep.results=TRUE
    )
    # remove duplicate models (bug in specr)
    sca= sca[!duplicated(sca[,c("x", "y", "controls")]),]
    scaest[i]= summarise_specs(sca)$median
    # "(2) Generate K different dependent variables under the null"
    yhat = apply(sca, 1, function(r) {
      t(data$y - r$res$coefficients[xnames]*data[xnames])
    })
    colnames(yhat) = paste("y", 1:ncol(yhat), sep="")
    data = cbind(data, yhat)
    medianperm= double(nperm)
    # Generate formulas for step 4 (using "predicted" y)
    formulas = apply(sca, 1, function(r) {r$res$call$formula})
    new_formulas = paste(
      colnames(yhat),
      sapply(formulas, function(f) {substring(f, 2)}, USE.NAMES=FALSE),
      sep=""
    )
    for (b in 1:length(medianperm)) {
      # "(3) Draw at random, and with replacement, N rows from this matrix, 
      #      using the same drawn rows of data for all K specifications."
      dataperm = data[sample(1:n, replace=TRUE),]
      # "(4) Estimate the K specifications on the drawn data."
      coefperm = sapply(new_formulas, function(f) {
        lm(f, dataperm)$coefficients[xnames]
      })
      medianperm[b] = median(coefperm)
    }
    # "(6) ... Compute what percentage..."
    scapval[i]= mean(abs(medianperm) > abs(scaest[i]))
}
```

We calculate the bias, RMSE and type I error rate.

```{r tab1, cache=TRUE}
bma.bias= mean(bmaest[,'x1'] - beta)
bma.rmse= sqrt(mean((bmaest[,'x1'] - beta)^2))
bma.reject= mean(margpp[,'x1']>0.95)
sca.bias= mean(scaest - beta)
sca.rmse= sqrt(mean((scaest - beta)^2))
sca.reject= mean(scapval< 0.05)
tab1= rbind(c(bma.bias, bma.rmse, bma.reject), c(sca.bias, sca.rmse, sca.reject))
rownames(tab1)= c('BMA','SCA')
colnames(tab1)= c('Bias','RMSE','Type I error')
tab1
```


## Truly non-zero treatment effect


```{r}
n= 1000; beta= 1; alpha= 1; nsims= 100
xnames= paste('x',1:length(beta),sep='')
znames= paste('z',1:length(alpha),sep='')
bmaest= margpp= matrix(NA,nrow=nsims,ncol=1+length(beta)+length(alpha))
colnames(bmaest)= colnames(margpp)= c('Intercept', xnames, znames)
scaest= double(nsims)
for (i in 1:nsims) {
    data= simdata(n=n,beta=beta,alpha=alpha,seed=100+i)
    #BMA
    ms = modelSelection(
      y=data[,1], x=data[,-1], verbose=FALSE, 
      priorCoef=zellnerprior(tau=nrow(data))
    )
    b= coef(ms)
    bmaest[i,]= b[-nrow(b),'estimate']
    margpp[i,]= b[-nrow(b),'margpp']
    #SCA
    sca= run_specs(df=data, y="y", x=xnames, controls=znames, model="lm")[-1,]
    scaest[i]= summarise_specs(sca)$median
}
```

When the coefficients are truly non-zero we calculate the bias, RMSE and power.

```{r tab2, cache=TRUE}
bma.bias= mean(bmaest[,'x1'] - beta)
bma.rmse= sqrt(mean((bmaest[,'x1'] - beta)^2))
bma.reject= mean(margpp[,'x1']>0.95)
sca.bias= mean(scaest - beta)
sca.rmse= sqrt(mean((scaest - beta)^2))
tab2= rbind(c(bma.bias, bma.rmse, bma.reject), c(sca.bias, sca.rmse, NA))
rownames(tab2)= c('BMA','SCA'); colnames(tab2)= c('Bias','RMSE','Power')
tab2
```



# Multiple treatments


## Zero average treatment effect

The simulation is as before, with the difference that `beta` has now 6 elements.

```{r}
n= 1000; beta= c(0,0,0,0,0,0); alpha= 1; nsims= 100
xnames= paste('x',1:length(beta),sep='')
znames= paste('z',1:length(alpha),sep='')
bmaest= margpp= matrix(NA,nrow=nsims,ncol=1+length(beta)+length(alpha))
colnames(bmaest)= colnames(margpp)= c('Intercept', xnames, znames)
scaest= double(nsims)
bma.ate= matrix(NA, nrow=nsims, ncol=2)
colnames(bma.ate)= c('estimate','margpp')
for (i in 1:nsims) {
    data= simdata(n=n,beta=beta,alpha=alpha,seed=200+i)
    #BMA
    ms = modelSelection(
      y=data[,1], x=data[,-1], verbose=FALSE, 
      priorCoef=zellnerprior(tau=nrow(data))
    )
    b = coef(ms)
    bmaest[i,]= b[-nrow(b),'estimate']
    margpp[i,]= b[-nrow(b),'margpp']
    #ATE
    bma.ate[i,]= getATE(ms, xvars=xnames, fun='mean')[c('estimate','margpp')]
    #SCA
    sca= run_specs(df=data, y="y", x=xnames, controls=znames, model="lm")
    scaest[i]= summarise_specs(unique(sca))$median
}
```

We first report the properties of individual coefficients.

```{r}
bias.indiv= rowMeans(t(bmaest[,xnames]) - beta)
rmse.indiv= sqrt(rowMeans((t(bmaest[,xnames]) - beta)^2))
reject.indiv= colSums(margpp[,xnames] > 0.95) / nrow(margpp)
tab3.indiv= rbind(bias.indiv, rmse.indiv, reject.indiv)
rownames(tab3.indiv)= c('Bias','RMSE','Type I error')
round(tab3.indiv, 5)
```

Next, we report the the properties of the ATE.

```{r tab3, cache=TRUE}
ate= mean(beta)
bma.biasate= mean(bma.ate[,'estimate'] - ate)
bma.rmseate= sqrt(mean((bma.ate[,'estimate'] - ate)^2))
bma.rejectate= mean(bma.ate[,'margpp']>0.95)

sca.biasate= mean(scaest - ate)
sca.rmseate= sqrt(mean((scaest - ate)^2))
tab3= rbind(c(bma.biasate, bma.rmseate, bma.rejectate), 
            c(sca.biasate, sca.rmseate, NA))
rownames(tab3)= c('BMA','SCA')
colnames(tab3)= c('Bias','RMSE','Type I error')
round(tab3, 4)
```



## Non-zero average treatment effect

We repeat the exercise with truly non-zero coefficinets.
Note: one can also use BMA to obtain inference on the median by using `getATE(ms, xvars=xnames, fun='median')` below.

```{r}
n= 1000; beta= c(0,0,1/4,3/4,1,1); alpha= 1; nsims= 100
xnames= paste('x',1:length(beta),sep='')
znames= paste('z',1:length(alpha),sep='')
bmaest= margpp= matrix(NA,nrow=nsims,ncol=1+length(beta)+length(alpha))
colnames(bmaest)= colnames(margpp)= c('Intercept', xnames, znames)
scaest= double(nsims)
bma.ate= matrix(NA, nrow=nsims, ncol=2)
colnames(bma.ate)= c('estimate','margpp')
for (i in 1:nsims) {
    data= simdata(n=n,beta=beta,alpha=alpha,seed=i)
    #BMA
    ms = modelSelection(
      y=data[,1], x=data[,-1], verbose=FALSE, 
      priorCoef=zellnerprior(tau=nrow(data))
    )
    b = coef(ms)
    bmaest[i,]= b[-nrow(b),'estimate']
    margpp[i,]= b[-nrow(b),'margpp']
    #ATE
    bma.ate[i,]= getATE(ms, xvars=xnames, fun='mean')[c('estimate','margpp')]
    #SCA
    sca= run_specs(df=data, y="y", x=xnames, controls=znames, model="lm")
    scaest[i]= summarise_specs(unique(sca))$median
}
```


```{r}
bias.indiv= rowMeans(t(bmaest[,xnames]) - beta)
rmse.indiv= sqrt(rowMeans((t(bmaest[,xnames]) - beta)^2))
reject.indiv= colSums(margpp[,xnames] > 0.95) / nrow(margpp)
tab4.indiv= rbind(bias.indiv, rmse.indiv, reject.indiv)
rownames(tab4.indiv)= c('Bias','RMSE','Proportion rejected')
round(tab4.indiv, 5)
```

```{r tab4, cache=TRUE}
ate= mean(beta)
bma.biasate= mean(bma.ate[,'estimate'] - ate)
bma.rmseate= sqrt(mean((bma.ate[,'estimate'] - ate)^2))
bma.rejectate= mean(bma.ate[,'margpp']>0.95)

sca.biasate= mean(scaest - ate)
sca.rmseate= sqrt(mean((scaest - ate)^2))
tab4= rbind(c(bma.biasate, bma.rmseate, bma.rejectate), 
            c(sca.biasate, sca.rmseate, NA))
rownames(tab4)= c('BMA','SCA')
colnames(tab4)= c('Bias','RMSE','Power')
round(tab4, 4)
```



```{r tabsummary, cache=TRUE, echo=FALSE}
# store the results for use in intro
nn= c('BMA bias =', 'RMSE =', 'type I error =', 'SCA bias =', 'RMSE =')
nn2= c('BMA bias =', 'RMSE =', 'power =', 'SCA bias =', 'RMSE =')

tab1summary= round(c(tab1[1,], tab1[2,-3]),4)
tab1summary= paste(nn,tab1summary)
tab2summary= round(c(tab2[1,], tab2[2,-3]),4)
tab2summary= paste(nn2,tab2summary)
tab3summary= round(c(tab3[1,], tab3[2,-3]),4)
tab3summary= paste(nn,tab3summary)
tab4summary= round(c(tab4[1,], tab4[2,-3]),4)
tab4summary= paste(nn2,tab4summary)
```



# Multiple outcomes

For the simulations with multiple outcomes, the function `simmultivdata` (created in the supplementary file `functions.R` ) generates the data $y$.

```{r}
L= 4; J= 5; q= 1; n= 1000; nsims= 100
Sigma= diag(L)
Sigma[1,2]= Sigma[1,3]= Sigma[2,3]= Sigma[2,1]= Sigma[3,1]= Sigma[3,2]= 0.9
Sigma[1:3,4]= 0.1; Sigma[4,1:3]= 0.1
```


## No average treatment effect

```{r}
beta= matrix(0,nrow=J,ncol=L); alpha= matrix(c(1,1,1,-1),nrow=q, ncol=L)

bma.ate= matrix(NA, nrow=nsims, ncol=2)
colnames(bma.ate)= c('estimate','margpp')
xnames= paste('x',1:J,sep='')
for (i in 1:nsims) {
    data= simmultivdata(n=n,beta=beta,alpha=alpha,Sigma=Sigma,seed=300+i)
    y= as.matrix(data[, 1:L])
    m= rowMeans(y)
    ms = modelSelection(
      y=m, x=data[,-1:-L], verbose=FALSE, 
      priorCoef=zellnerprior(tau=nrow(data))
    )
    bma.ate[i,]= getATE(ms, xvars=xnames)[c('estimate','margpp')]
}
```

```{r}
ate= mean(beta)
bma.biasate= mean(bma.ate[,'estimate'] - ate)
bma.rmseate= sqrt(mean((bma.ate[,'estimate'] - ate)^2))
bma.rejectate= mean(bma.ate[,'margpp']>0.95)
```

```{r}
tab5= matrix(c(bma.biasate, bma.rmseate, bma.rejectate), nrow=1)
colnames(tab5)= rbind(c('Bias','RMSE','Type I error'))
rownames(tab5) = c('BMA')
round(tab5, 4)
```


## Non-zero average treatment effect

```{r}
beta= matrix(0,nrow=J,ncol=L)
beta[,1]= beta[,2]= beta[,3]= c(1,1,1,0,0)
beta[,4]= c(1/4, 1/4, 1/4, 0, 0)
alpha= matrix(c(1,1,1,-1),nrow=q, ncol=L)
```

```{r}
bma.ate= matrix(NA, nrow=nsims, ncol=2)
colnames(bma.ate)= c('estimate','margpp')
xnames= paste('x',1:J,sep='')
for (i in 1:nsims) {
    data= simmultivdata(n=n,beta=beta,alpha=alpha,Sigma=Sigma,seed=300+i)
    y= as.matrix(data[, 1:L])
    m= rowMeans(y)
    ms = modelSelection(
      y=m, x=data[,-1:-L], verbose=FALSE, 
      priorCoef=zellnerprior(tau=nrow(data))
    )
    bma.ate[i,]= getATE(ms, xvars=xnames)[c('estimate','margpp')]
}
```

```{r}
ate= mean(beta)
bma.biasate= mean(bma.ate[,'estimate'] - ate)
bma.rmseate= sqrt(mean((bma.ate[,'estimate'] - ate)^2))
bma.rejectate= mean(bma.ate[,'margpp']>0.95)
```

```{r}
tab6= matrix(c(bma.biasate, bma.rmseate, bma.rejectate), nrow=1)
colnames(tab6)= c('Bias','RMSE','Power')
rownames(tab6) = c('BMA')
round(tab6, 4)
```

# Export tables

Store all the main results in a summary table.

```{r}
l = list(tab1, tab2, tab3, tab4, tab5, tab6)
tab = bind_rows(
  lapply(l, function(m) {
    bind_cols(rownames(m), as.data.frame(m))
  })
)
tab = as.data.frame(tab) %>%
  as.data.frame(row.names = 1:nrow(.)) %>%
  subset(select = -c(`Type I error`, Power)) %>%
  rename(Estimator = "...1") %>%
  add_column(
    Scenario = c(
      '', '\\multirow{-2}*{$\\beta = (0)$}',
      '', '\\multirow{-2}*{$\\beta = (1)$}',
      '', '\\multirow{-2}*{$\\beta = (0, 0, 0, 0, 0, 0)$}',
      '', '\\multirow{-2}*{$\\beta = (0, 0, 0.25, 0.75, 1, 1)$}',
      'GATE = 0',
      'GATE $\\neq$ 0'
    ),
    .before = "Estimator"
  ) %>%
  mutate(id=row_number()) %>% 
  tibble::column_to_rownames('id')
```

Export the tables to latex.

```{r}
print(
  xtable(
    tab, 
    digits = 3,
    label = "tab:sim_summary",
    caption = "Bias and root mean squared error in simulation scenarios 1-5. Scenarios 1-2 have a single treatment, Scenarios 3-4 have 6 treatments, and Scenarios 5-6 have 4 treatments and 5 outcomes. $\\beta$ indicates the data-generating truth."
  ),
  comment = FALSE,
  sanitize.text.function = force,
  file = file.path(PATH, 'data/export/sim_summary.tex'),
)

print(
  xtable(
    tab3.indiv, 
    digits = 3,
    label = "tab:sim_3_indiv",
    caption = "Individual coefficient results for BMA in simulation \\#3 with coefficients $\\beta = (0, 0, 0, 0, 0, 0)$"
  ),
  comment = FALSE,
  sanitize.text.function = force,
  file = file.path(PATH, 'data/export/sim_3_indiv.tex'),
)

print(
  xtable(
    tab4.indiv, 
    digits = 3,
    label = "tab:sim_4_indiv",
    caption = "Individual coefficient results for BSCA in simulation \\#4 with coefficients $\\beta = (0, 0, 0.25, 0.75, 1, 1)$"
  ),
  comment = FALSE,
  sanitize.text.function = force,
  file = file.path(PATH, 'data/export/sim_4_indiv.tex'),
)
```

